Minimizing Time Required For Graduation

### 2010-12-05 paper/paper.tex
fixed bib errors, number equations

To solve this issues, we introduce a set of rules for each course. Each rule is
a boolean expression represented in propositional logic in disjunctive normal
form. For example, the rule associated with $MATH327$ is $$ \text{{\it MATH327:
(MATH126 AND MATH300) OR MATH136}}.  $$ As previously stated, we denote the set
of all rules for all classes as $P$.

@@

With these functions defined, we can now define our heuristic function:
\begin{equation} Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c,
q) + \gamma * Quarters(c) \end{equation} Where $c$ is a given course offering,
$q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real
numbers, giving a weight to piece of the heuristic. The scalars are determined
experimentally.

\subsubsection{Formulation as Binary Integer Program} We can now formulate our
problem as a binary linear programming problem. Before stating the linear
program, we must first define some variables used in it: \\ \begin{itemize}
\item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
we don't take } X_i \end{array} \right. $ \item For all $t \in O$, we let $t
= 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our problem as
follows:\\ Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
$$ Subject To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ \forall t_j \in
O, \sum_{i=1}^m t_j * X_i \leq 1, $$ $$ X_i \in \{0, 1\} $$ The linear program
can be interpreted as follows: \begin{itemize} \item The value of a quarter is
the total number of courses taken that quarter.  \item In each quarter, the
total number of credits cannot exceed the specified limit.  \item Only one
course can be taken in one particular time slot \item A course can only be taken
or not taken. In other words, decision variables are binary.  \item Each
non-zero decision variable in the optimal solution represents a class we will
take.  \end{itemize}

It is worth nothing that all of the t variables are essentially dummy variables
that ensure the correct number of equations of type 2. Each of the equations
based on eqn. 2 ensure that we only take one class per time period per day. Also
this linear program is ‘nice’ in that it always has feasible origin (don’t take
any classes), and that the simplest formulation is already in standard form.

@@

We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding
dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the
graph. To ensure that $v'$ does not alter the final optimal solution, we assign
it the following properties: \begin{itemize} \item $Weight(v') = l+1$ \item
$Value(v') = 0$ \end{itemize} As the weight of our dummy node is greater than
our credit limit, and it's value is zero, we know that it will never appear in
the optimal solution. With the above reasoning we show for our case of class
schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding
a dummy node without loss of generality holds. We can now conveniently write the
class picking problem as the knapsack with conflict graphs problem (KCG): \\
Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i $$ Subject
To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ X_i + X_j \leq 1 \text{ for
all connected vertices } i, j \in G_{conflict}, $$ $$ X_i \in \{0, 1\} $$ It is
worth noting that for relatively small versions of this problem, in the low
hundreds of variables it is possible to solve this problem exactly using modern
LP solving software \cite{yamada:heuristic}. In our case, just solving for the
optimal classes for a single major fits well into the scope of a few hundred (or
even just tens of) variables. For multiple majors or to consider class selection
for distribution and elective requirements, however the number of classes
balloons significantly and the KCG problem becomes much harder. For this version
of the problem we suggest using the techniques used by Yamada et al
\cite{yamada:heuristic}, which we briefly outline below.

>>>>

To solve this issues, we introduce a set of rules for each course. Each rule is
a boolean expression represented in propositional logic in disjunctive normal
form. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.  $$ As previously
stated, we denote the set of all rules for all classes as $P$.

@@

With these functions defined, we can now define our heuristic function:

\begin{equation} 
    Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c,q) 
    - \gamma * Quarters(c)
    \label{value_func}
\end{equation}

Where $c$ is a given course offering,
$q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real
numbers, giving a weight to piece of the heuristic. The scalars are determined
experimentally.

\subsubsection{Formulation as Binary Integer Program} We can now formulate our
problem as a binary linear programming problem. Before stating the linear
program, we must first define some variables used in it: \\ \begin{itemize}
\item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
we don't take } X_i \end{array} \right. $ \item For all $t \in O$, we let $t
= 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our problem as
follows:

Maximize:

\begin{equation}
    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
    \label{qtr_val}
\end{equation}

Subject To:

\begin{equation}
    \sum_{i=1}^m Weight(c_i) * X_i \leq l, \forall t_j \in O,
    \label{weight_lim}
\end{equation}

\begin{equation}
    \sum_{i=1}^m t_j * X_i \leq 1, X_i \in \{0, 1\}
    \label{time_con}
\end{equation}

The linear program
can be interpreted as follows: \begin{itemize} \item The value of a quarter is
the total number of courses taken that quarter.  \item In each quarter, the
total number of credits cannot exceed the specified limit.  \item Only one
course can be taken in one particular time slot \item A course can only be taken
or not taken. In other words, decision variables are binary.  \item Each
non-zero decision variable in the optimal solution represents a class we will
take.  \end{itemize}

Each of the equations based on eqn. 2 ensure that we only take one class per 
time period per day. Also this linear program is ‘nice’ in that it always has 
feasible origin (don’t take any classes), and that the simplest formulation 
is already in standard form.

@@

We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding
dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the
graph. To ensure that $v'$ does not alter the final optimal solution, we assign
it the following properties: \begin{itemize} \item $Weight(v') = l+1$ \item
$Value(v') = 0$ \end{itemize} As the weight of our dummy node is greater than
our credit limit, and it's value is zero, we know that it will never appear in
the optimal solution. With the above reasoning we show for our case of class
schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding
a dummy node without loss of generality holds. We can now conveniently write the
class picking problem as the knapsack with conflict graphs problem (KCG):

Maximize:

\begin{equation}
    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
\end{equation}

Subject To:

\begin{equation}
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
\end{equation}

\begin{equation}
X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
\end{equation}

\begin{equation}
X_i \in \{0, 1\}
\end{equation}

For versions of this problem in the low hundreds of variables it is possible 
to solve this problem exactly using modern LP solving software 
\cite{yamada:heuristic}. In our case, just solving for the optimal classes for 
a single major fits well into the scope of a few hundred (or even just tens of) 
variables. For multiple majors or to consider class selection
for distribution and elective requirements, however the number of classes
balloons significantly and the KCG problem becomes much harder. For this version
of the problem we suggest using the techniques used by Yamada et al
\cite{yamada:heuristic}, which we briefly outline below.
